BANTER for Beaked Whales: South Atlantic

### 2023-02-07 satlantic/beakerBanter_satlantic.Rmd
Update manuscript and hawaii data
Extensive work on Manuscript and preparation of Hawaii data (in folder). Hawaii data *should* be good to knit. Manuscript has references, needs updated results, figures, inline data, and updated discussion.

---
title: 'BANTER for Beaked Whales: North Atlantic'
author: "Shannon Rankin"
date: "2023-01-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

@@

1.  **Set up our PPS** (PAMPal Settings Object) for the NAtlantic
    dataset

@@

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data <- readRDS('natlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

@@

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='natlantic_bw')
saveRDS(data, 'natlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

@@

```{r, assign species ID}
HB1603_BW <- read_excel("HB1603_BWs.xlsx", sheet = "Alldata")#read in data
#create formatted event/species dataframe 
speciesID <- HB1603_BW[c("PGId", "Final Species Classification", "Database")] %>% 
  mutate("event" = paste(Database, PGId, sep=".OE")) %>% 
  rename("species" = "Final Species Classification") %>% 
  mutate("species" = str_replace_all(species, "'", ""),
         "species" = str_replace(species, "Gervais/Trues", "BW")) %>% 
  select(all_of(c("event", "species")))

@@

```{r, filter unwanted data}
goodSpecies<- c(c("Trues", "Cuviers", "Sowerbys", "Gervais"))
data <- filter(data, species %in% goodSpecies)
```

@@

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data, dropSpecies = 'unid', 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='natlantic_banter_data.rds')

banter_data_ici <- export_banter(data, dropSpecies = 'unid', training=TRUE)
saveRDS(banter_data_ici, file='natlantic_banter_data_ici.rds')
```

@@

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model <- readRDS('natlantic_banter_model_t5e3s3_t5e3s4.rds')
```

```{r, banter detector model, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=5e3, sampsize=3, importance = TRUE)
```

@@

```{r,save banter model, eval=freshRun}
saveRDS(banter_model, 'natlantic_banter_model_t5e3s3_t5e3s4.rds')
```

@@

```{r, read banter ici model, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('natlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

@@

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'natlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

@@

```{r, confusion matrix, include=TRUE}
natlantic_confuseMatrix <- rfPermute::confusionMatrix(banter_model_RF)
natlantic_confuseMatrix <- kable(natlantic_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname, "natlantic_confuseMatrix.png", sep="_"), zoom = 2)

natlantic_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
natlantic_ici_confuseMatrix <- kable(natlantic_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname_ici, "natlantic_ici_confuseMatrix.png", sep="_"), zoom = 2)

@@

![](banter_model_natlantic_confuseMatrix.png)

@@

```{r, proximity plot, include=TRUE}
png(paste(modelname, "natlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
natlantic_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
natlantic_proximityPlot <- plotProximity(banter_model_RF)

png(paste(modelname_ici, "natlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
ici_natlantic_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_natlantic_proximityPlot <- plotProximity(banter_model_ici_RF)
```

@@

```{r, importance heat map, include=TRUE}
png(paste(modelname, "natlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
natlantic_importance <- plotImportance(banter_model_RF, plot.type="heatmap")

png(paste(modelname_ici, "natlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_natlantic_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

@@

```{r, plot votes, include=TRUE}
png(paste(modelname, "natlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
natlantic_votes <- plotVotes(banter_model_RF)

png(paste(modelname_ici, "natlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_natlantic_votes <- plotVotes(banter_model_ici_RF)
```

@@

```{r, class priors, include=TRUE}
natlantic_priors <- classPriors(banter_model_RF, NULL)
```

>>>>

---
title: 'BANTER for Beaked Whales: South Atlantic'
author: "Shannon Rankin"
date: "2023-02-03"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

@@

1.  **Set up our PPS** (PAMPal Settings Object) for the satlantic
    dataset

@@

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data <- readRDS('satlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

@@

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='satlantic_bw')
saveRDS(data, 'satlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

@@

```{r, assign species ID}
GU1605_BW <- read_excel("GU1605_PG_OfflineEvents_20210809.xlsx", sheet = "AllData")#read in data
#create formatted event/species dataframe 
speciesID <- GU1605_BW[c("Id", "Species", "Database")] %>% 
  mutate("event" = paste(Database, Id, sep=".OE")) %>% 
  rename("species" = "Species") %>% 
  mutate("species" = str_replace_all(species, "'", ""),
         "species" = str_replace(species, "Sowerby", "Sowerbys"),
         "species" = str_replace(species, "Blainville", "Blainvilles"),
         "species" = str_replace(species, "Cuvier", "Cuviers")) %>% 
  select(all_of(c("event", "species")))

@@

```{r, filter unwanted data}
goodSpecies<- c(c("Blainvilles", "Trues", "Cuviers", "Sowerbys", "Gervais"))
data <- filter(data, species %in% goodSpecies)
```

@@

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data, dropSpecies = 'unid', 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='satlantic_banter_data.rds')

banter_data_ici <- export_banter(data, dropSpecies = 'unid', training=TRUE)
saveRDS(banter_data_ici, file='satlantic_banter_data_ici.rds')
```

@@

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model <- readRDS('satlantic_banter_model_t5e3s3_t5e3s4.rds')
```

```{r, banter detector model, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=5e3, sampsize=4, importance = TRUE)
```

@@

```{r,save banter model, eval=freshRun}
saveRDS(banter_model, 'satlantic_banter_model_t5e3s3_t5e3s4.rds')
```

@@

```{r, read banter ici model, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('satlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

@@

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'satlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

@@

```{r, confusion matrix, include=TRUE}
satlantic_confuseMatrix <- rfPermute::confusionMatrix(banter_model_RF)
satlantic_confuseMatrix <- kable(satlantic_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname, "satlantic_confuseMatrix.png", sep="_"), zoom = 2)

satlantic_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
satlantic_ici_confuseMatrix <- kable(satlantic_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname_ici, "satlantic_ici_confuseMatrix.png", sep="_"), zoom = 2)

@@

![](banter_model_hawaii_confuseMatrix.png)

@@

```{r, proximity plot, include=TRUE}
png(paste(modelname, "satlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
hawaii_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
hawaii_proximityPlot <- plotProximity(banter_model_RF)

png(paste(modelname_ici, "satlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
```

@@

```{r, importance heat map, include=TRUE}
png(paste(modelname, "satlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
hawaii_importance <- plotImportance(banter_model_RF, plot.type="heatmap")

png(paste(modelname_ici, "satlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_hawaii_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

@@

```{r, plot votes, include=TRUE}
png(paste(modelname, "satlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
hawaii_votes <- plotVotes(banter_model_RF)

png(paste(modelname_ici, "satlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_hawaii_votes <- plotVotes(banter_model_ici_RF)
```

@@

```{r, class priors, include=TRUE}
hawaii_priors <- classPriors(banter_model_RF, NULL)
```
